### 2020-12-10 paper.tex
contrasting weak and strong fulfillment

\RequirePackage{amssymb}  %% bizarre error previous def of Bbbk if after documentclass
\documentclass[acmsmall,screen]{acmart}\settopmatter{printfolios=true}
\hypersetup{bookmarksnumbered,bookmarksopen=true,bookmarksdepth=3}
\settopmatter{printfolios=true}
\AtEndPreamble{%
  % \theoremstyle{acmdefinition}
  % \theoremstyle{plain}
  % \newtheorem{theorem}{Theorem}[section]
  % \newtheorem{proposition}[theorem]{Proposition}
  % \theoremstyle{acmdefinition}
  % \newtheorem{definition}[theorem]{Definition}
  \theoremstyle{acmdefinition}
  \newtheorem{remark}[theorem]{Remark}
  \newtheorem{candidate}[theorem]{Candidate}
  \renewcommand{\theequation}{\fnsymbol{equation}}
}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations

% \setcopyright{rightsretained}
% \acmPrice{?}
% \acmDOI{?}
% \acmYear{2021}
% \copyrightyear{2021}
% \acmSubmissionID{1}
\acmJournal{PACMPL}
% \acmVolume{1}
% \acmNumber{POPL}
% \acmArticle{1}
% \acmMonth{1}
% \startPage{1}

\usepackage{macros}
\begin{document}

\title{Weak Pomsets}
% \author{Radha Jagadeesan}
% \orcid{0000-0002-4525-1589}
% \affiliation{
%   %\department{School of Computing}
%   \institution{DePaul University}
%   \city{Chicago}
%   \country{USA}
% }
% %\email{rjagadeesan@cs.depaul.edu}

% \author{Alan Jeffrey}
% \orcid{0000-0001-6342-0318}
% \affiliation{
%   \institution{Mozilla Research and the Servo Project}
%   \city{Chicago}
%   \country{USA}
% }
% %\email{ajeffrey@mozilla.com}

% \author{James Riely}
% \orcid{0000-0002-8731-1463}
% \affiliation{
%   %\department{School of Computing}
%   \institution{DePaul University}
%   \city{Chicago}
%   \country{USA}
% }
% %\email{jriely@cs.depaul.edu}

%\maketitle
\input{model}
\clearpage
\bibliography{bib}
\appendix
\end{document}

% Local Variables:
% mode: latex
% TeX-master: t
% End: