### 2021-06-09 paper.tex
added some elimination examples

% POPL
% Thu 8 Jul 2021  Submission Deadline
% Sat 4 - Wed 8 Sep 2021 Author Response Period
% Fri 1 Oct 2021 Notification

\RequirePackage{amssymb}  %% bizarre error previous def of Bbbk if after documentclass
\documentclass[acmsmall,screen,anonymous,review]{acmart}\settopmatter{printfolios=true}
\hypersetup{bookmarksnumbered,bookmarksopen=true,bookmarksdepth=3}
\settopmatter{printfolios=true}
\AtEndPreamble{%
  % \theoremstyle{acmdefinition}
  % \theoremstyle{plain}
  % \newtheorem{theorem}{Theorem}[section]
  % \newtheorem{proposition}[theorem]{Proposition}
  % \theoremstyle{acmdefinition}
  % \newtheorem{definition}[theorem]{Definition}
  \theoremstyle{acmdefinition}
  \newtheorem{remark}[theorem]{Remark}
  \newtheorem{candidate}[theorem]{Candidate}
  \renewcommand{\theequation}{\fnsymbol{equation}}
}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations

\setcopyright{rightsretained}
\acmPrice{}
\acmDOI{}
\acmYear{2022}
\copyrightyear{2022}
\acmSubmissionID{}
\acmJournal{PACMPL}
\acmVolume{0}
\acmNumber{POPL}
\acmArticle{0}
\acmMonth{1}
\startPage{1}

\usepackage{macros}
% \makeatletter
% % get ams to ignore tag in spacing
% \renewcommand{\tagform@}[1]{\maketag@@@{\makebox[0pt][r]{(\ignorespaces#1\unskip\@@italiccorr)}}}
% \newcommand{\spacetag}[2][3em]{\makebox[#1]{}\tag{#2}}
% \newcommand{\notagparens}{\renewcommand{\tagform@}[1]{\maketag@@@{\makebox[0pt][r]{\ignorespaces##1\unskip\@@italiccorr}}}}
% \makeatother

\begin{document}

\title{A Unified Memory Model for Heterogenous Systems}

% \author{James Riely}
% \orcid{0000-0002-8731-1463}
% \affiliation{
%   %\department{School of Computing}
%   \institution{DePaul University}
%   \city{Chicago}
%   \country{USA}
% }
% %\email{jriely@cs.depaul.edu}

\maketitle
%\input{intro.tex}
\input{definitions.tex}
%\input{drf.tex}
%\input{model.tex}
%\input{popo.tex}
\clearpage
\bibliography{bib}
\appendix
\end{document}

% Local Variables:
% mode: latex
% TeX-master: t
% End: